---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# abstr


<!-- badges: start -->
[![R-CMD-check](https://github.com/a-b-street/abstr/workflows/R-CMD-check/badge.svg)](https://github.com/a-b-street/abstr/actions)
<!-- badges: end -->
  
The goal of abstr is to provide an R interface to the [A/B Street](https://github.com/a-b-street/abstreet#ab-street) transport planning/simulation game.
Currently it provides a way to convert aggregated origin-destination data, combined with data on buildings representing origin and destination locations, into `.json` files that can be directly imported into the A/B Street game.
See https://a-b-street.github.io/docs/dev/formats/scenarios.html#example for details of the schema that the package outputs.

## Installation

You can install the released version of abstr from
<!-- [CRAN](https://CRAN.R-project.org) with: -->
GitHub as follows:

```r
remotes::install_github("a-b-street/abstr")
```

## Example

```{r output-sf, message=FALSE, warning=FALSE}
library(abstr)
leeds_desire_lines # origin destination data with aggregate travel data
plot(leeds_desire_lines)
ablines = ab_scenario(
 leeds_houses,
 leeds_buildings,
 dslines,
 leeds_zones,
 output_format = "sf"
)
plot(dslines$geometry, lwd = dslines[[3]] / 30)
plot(leeds_site_area$geometry, add = TRUE)
plot(leeds_buildings$geometry, add = TRUE)
plot(ablines$geometry, col = "blue", add = TRUE)
```

Each blue line in the plot above represents a single trip, with an associated depature time, that can be represented in A/B Street.

You can output the result as a list object that can be saved as a JSON file as follows, taking only one of the desire lines (desire line 7, which has only 9 trips for ease of viewing the results) as an example:


```{r output-json, message=FALSE, warning=FALSE}
library(abstr)
ab_scenario_list = ab_scenario(
 leeds_houses,
 leeds_buildings,
 leeds_desire_lines[7, ],
 leeds_zones,
 output_format = "json_list"
)
ab_scenario_list
ab_save(ab_scenario_list, "ab_scenario.json")
```

Let's see what is in the file:

```r
file.edit("ab_scenario.json")
```

It should look something like this, matching [A/B Street's schema](https://a-b-street.github.io/docs/dev/formats/scenarios.html#example).

```json
{
  "scenario_name": "baseline",
  "people": [
    {
      "origin": {
        "Position": {
          "longitude": -1.518,
          "latitude": 53.7896
        }
      },
      "trips": [
        {
          "departure": 27967,
          "destination": {
            "Position": {
              "longitude": -1.5435,
              "latitude": 53.7793
            }
          },
          "mode": "other"
        }
      ]
    },
```
