---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# abstr


<!-- badges: start -->
[![R-CMD-check](https://github.com/Robinlovelace/abstr/workflows/R-CMD-check/badge.svg)](https://github.com/Robinlovelace/abstr/actions)
<!-- badges: end -->
  
The goal of abstr is to provide an R interface to the A/B Street transport planning/simulation game.
In the first instance, it provides a way to generate scenarios of change and saving them as `.json` files that can be directly imported into the A/B Street game.
See https://a-b-street.github.io/docs/dev/formats/scenarios.html#example for details of the schema that the package outputs.

## Installation

You can install the released version of abstr from
<!-- [CRAN](https://CRAN.R-project.org) with: -->
GitHub as follows:

```r
remotes::install_github("cyipt/abstr")
```

## Example

```{r output-sf, message=FALSE, warning=FALSE}
library(abstr)
dslines = leeds_desire_lines
ablines = ab_scenario(
 leeds_houses,
 leeds_buildings,
 dslines,
 leeds_zones,
 output_format = "sf"
)
plot(dslines$geometry, lwd = dslines[[3]] / 30)
plot(leeds_site_area$geometry, add = TRUE)
plot(leeds_buildings$geometry, add = TRUE)
plot(ablines$geometry, col = "blue", add = TRUE)
```

