---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# abstr


<!-- badges: start -->
[![R-CMD-check](https://github.com/a-b-street/abstr/workflows/R-CMD-check/badge.svg)](https://github.com/a-b-street/abstr/actions)
<!-- badges: end -->
  
The goal of abstr is to provide an R interface to the [A/B Street](https://github.com/a-b-street/abstreet#ab-street) transport planning/simulation game.
Currently it provides a way to convert aggregated origin-destination data, combined with data on buildings representing origin and destination locations, into `.json` files that can be directly imported into the A/B Street game.
See https://a-b-street.github.io/docs/dev/formats/scenarios.html#example for details of the schema that the package outputs.

## Installation

You can install the released version of abstr from
<!-- [CRAN](https://CRAN.R-project.org) with: -->
GitHub as follows:

```r
remotes::install_github("a-b-street/abstr")
```

## Example

The example below shows how `abstr` can be used.
The input datasets include `sf` objects representing houses, buildings, origin-destination (OD) data represented as desire lines and administrative zones representing the areas within which trips in the desire lines start and end.
With the exception of OD data, each of the input datasets is readily available for most cities.
The input datasets are illustrated in the plots below, which so a small example dataset taken from the city of Leeds, UK.

```{r input, fig.cap="Example data that can be used as an input by functions in abstr to generate trip-level scenarios that can be imported by A/B Street."}
library(abstr)
library(tmap) # for map making
tm_shape(leeds_zones) + tm_polygons(col = "grey") +
  tm_shape(leeds_site_area) + tm_polygons(col = "red") +
  tm_shape(leeds_houses) + tm_polygons(col = "yellow") +
  tm_shape(leeds_buildings) + tm_polygons(col = "blue") +
  tm_shape(leeds_desire_lines) + tm_lines(lwd = "all_base", scale = 3)
```



```{r output-sf, message=FALSE, warning=FALSE}
ablines = ab_scenario(
 houses = leeds_houses,
 buildings = leeds_buildings,
 desire_lines = leeds_desire_lines,
 zones = leeds_zones,
 output_format = "sf"
)
tmap_mode("view")
bb = tmaptools::bb(leeds_houses, 10)
tm_shape(leeds_buildings, bbox = bb) + tm_polygons() +
  tm_shape(leeds_houses) + tm_polygons(col = "blue") +
  tm_shape(ablines) + tm_lines(col = "mode_base") 
```

Each line in the plot above represents a single trip, color representing mode.
Each trip has an associated departure time, that can be represented in A/B Street.

Under a different scenario, the Go Dutch scenario of active travel uptake represented in the columns containing `godutch` for example, the travel patterns would be substantially different.
In the aggregated desire lines, the differences between the two scenarios are substantial, as shown in the table below:

```{r scentable}
desire_line_data = sf::st_drop_geometry(leeds_desire_lines)
nms = names(desire_line_data)
nms
nms_scenarios = nms[grepl(pattern = "base|dutch", x = nms)]
knitr::kable(desire_line_data[nms_scenarios])
```

The Go Dutch scenario can be disaggregated so that trips start and begin in buildings, as shown below.

```{r abdutch}
ablines_dutch = ab_scenario(
 houses = leeds_houses,
 buildings = leeds_buildings,
 desire_lines = leeds_desire_lines,
 zones = leeds_zones,
 output_format = "sf"
)
tm_shape(leeds_buildings, bbox = bb) + tm_polygons() +
  tm_shape(leeds_houses) + tm_polygons(col = "blue") +
  tm_shape(ablines_dutch) + tm_lines(col = "mode_base") 
```

<!-- todo: add time to df -->

You can output the result as a list object that can be saved as a JSON file as follows, taking only one of the desire lines (desire line 7, which has only 9 trips for ease of viewing the results) as an example:

```{r output-json, message=FALSE, warning=FALSE}
library(abstr)
ab_scenario_list = ab_scenario(
 leeds_houses,
 leeds_buildings,
 leeds_desire_lines,
 leeds_zones,
 output_format = "json_list"
)
ab_scenario_list
ab_save(ab_scenario_list, "ab_scenario.json")
```

Let's see what is in the file:

```r
file.edit("ab_scenario.json")
```

The first trip schedule should look something like this, matching [A/B Street's schema](https://a-b-street.github.io/docs/dev/formats/scenarios.html#example).

```json
{
  "scenario_name": "base",
  "people": [
    {
      "origin": {
        "Position": {
          "longitude": -1.5278,
          "latitude": 53.7888
        }
      },
      "trips": [
        {
          "departure": 28236,
          "destination": {
            "Position": {
              "longitude": -1.5717,
              "latitude": 53.8039
            }
          },
          "mode": "Walk"
        }
      ]
    }
```
