---
title: "Get started with abstr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with abstr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(abstr)
```

## Getting started

In order to import scenario files into A/B Street, you will need to:

-   Install a stable version of [Rstudio and R](https://www.rstudio.com/products/rstudio/)

-   Install a stable version of [Rust](https://www.rust-lang.org/tools/install)

    -   On Windows, you will also need [Visual Code Studio](https://code.visualstudio.com/) and [Visual Studio c++ build tools](https://visualstudio.microsoft.com/downloads/) prior to installing Rust.

-   On Linux, run `sudo apt-get install libasound2-dev libxcb-shape0-dev libxcb-xfixes0-dev libpango1.0-dev libgtk-3-dev` or the equivalent for your distribution.

-   Download the A/B Street repo `git clone https://github.com/a-b-street/abstreet.git`

-   Fetch the minimal amount of data needed to get started `cargo run --bin updater -- --minimal`

# Projects supporting `abstr`

Several open source R packages enabled the creation of `abstr`.

These include:

-   [jsonlite](https://cran.r-project.org/web/packages/jsonlite/index.html) for awesome JSON parsing and generation
-   [magrittr](https://cran.r-project.org/web/packages/magrittr/index.html) for magic `%>%` pipe-command chaining
-   [sf](https://cran.r-project.org/web/packages/sf/index.html) for simple feature support
-   [tibble](https://cran.r-project.org/web/packages/tibble/index.html) for strict data-frame declaration
-   [tidyr](https://cran.r-project.org/web/packages/tidyr/index.html) for all things data cleaning

For more on the history motivating the development of the `abstr` package see the [`pct_to_abstr` vignette]()
