---
title: "Reproducing Montlake Eastside Seattle, US"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing Montlake Eastside Seattle, US}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
author: 
- "Robin Lovelace, Trevor Nederlof and Nathanael Sheehan"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction 
Lets start with the city A/B street began, Seattle. U.S. The seaport city is home to over 700,000 people, including A/B Street creator, Dustin Carlino, who has been developing tools to empirically study the impact of small changes within the road network: this means "you can transform that street parking into a bus lane or fix that pesky left turn at a traffic signal, measure the effects, then propose actually making the change". For the past two years, Seattle has been a key are of study within the A/B street simuverse and thus makes it a great starting point in understanding the utilities of `abstr` to generate site data for A/B street. 

<!-- ## Install packages -->
<!-- ```{r, message=FALSE,warning=FALSE} -->
<!-- devtools::install_cran(tidyverse) -->
<!-- ``` -->

## Example


### Load packages
```{r, message=FALSE,warning=FALSE}
library(tidyverse)
library(sf)
library(abstr)
```


### Fetch Montlake Polygon
```{r, message=FALSE,warning=FALSE}
montlake_poly_url <- "https://raw.githubusercontent.com/a-b-street/abstreet/master/importer/config/us/seattle/montlake.poly"

raw_boundary_vec <- readr::read_lines(montlake_poly_url)
boundary_matrix <- raw_boundary_vec[(raw_boundary_vec != "boundary") & (raw_boundary_vec != "1") & (raw_boundary_vec != "END")] %>%
  stringr::str_trim() %>%
  tibble::as_tibble() %>%
  dplyr::mutate(y_boundary = as.numeric(lapply(stringr::str_split(value, "    "), `[[`, 1)),
                x_boundary = as.numeric(lapply(stringr::str_split(value, "    "), `[[`, 2))) %>%
  dplyr::select(-value) %>%
  as.matrix()
boundary_sf_poly <- sf::st_sf(geometry = sf::st_sfc(sf::st_polygon(list(boundary_matrix)), crs = 4326))

```

### Filter Polygon by Zones
```{r, message=FALSE,warning=FALSE}
all_zones_tbl <- sf::st_read("https://raw.githubusercontent.com/psrc/soundcast/master/inputs/base_year/taz2010.geojson") %>% sf::st_transform(4326)
zones_in_boundary_tbl <- all_zones_tbl[sf::st_intersects(all_zones_tbl, boundary_sf_poly, sparse = F),]
```


### Generate OD Matrix
```{r, message=FALSE,warning=FALSE}
## create a OD matrix
od_tbl_long <- dplyr::select(all_trips_tbl, otaz, dtaz, mode) %>%
  dplyr::mutate(mode = dplyr::case_when(mode %in% c(1, 9) ~ "Walk",
                                        mode == 2 ~ "Bike",
                                        mode %in% c(3, 4, 5) ~ "Drive",
                                        mode %in% c(6, 7, 8) ~ "Transit",
                                        TRUE ~ as.character(NA))) %>%
  dplyr::filter(!is.na(mode)) %>%
  dplyr::group_by(otaz, dtaz, mode) %>%
  dplyr::summarize(n = n()) %>%
  dplyr::ungroup() %>%
  # only keep an entry if the origin or destination is in a Montlake zone
  dplyr::filter((otaz %in% zones_in_boundary_tbl$TAZ) | (dtaz %in% zones_in_boundary_tbl$TAZ))

# create a wide OD matrix and filter out any OD entries with under 25 trips in it
montlake_od_tbl <- tidyr::pivot_wider(od_tbl_long, names_from = mode, values_from = n, values_fill = 0) %>%
  dplyr::rename(o_id = otaz, d_id = dtaz) %>%
  dplyr::mutate(total = Drive + Transit + Bike + Walk) %>%
  dplyr::filter(total >= 25) %>%
  dplyr::select(-total)

montlake_zone_tbl <- dplyr::right_join(all_zones_tbl,
                                       tibble::tibble("TAZ" = unique(c(montlake_od_tbl$o_id, montlake_od_tbl$d_id))),
                                       by = "TAZ") %>%
  dplyr::select(TAZ) %>%
  dplyr::rename(id = TAZ)
```
